---
title: "Conditional Probability"
author: "Kaiona Apio"
date: "02/17/2025"

format: 
  html: 
    theme: superhero  
    mainfont: monospace
    highlight-style: github
    title-block-banner: true
    embed-resources: true
---

# 1. Setup

**Step Up Code:**

```{r, warning=FALSE}
sh <- suppressPackageStartupMessages
sh(library(tidyverse))
sh(library(caret))
sh(library(tidytext))
wine <- readRDS(gzcon(url("https://github.com/cd-public/D505/raw/master/dat/pinot.rds")))
```

# 2. Conditional Probability

Calculate the probability that a Pinot comes from Burgundy given it has the word 'fruit' in the description.

```{r, Feature Engineering}
wino <- wine%>%
  mutate(fruit = str_detect(description, "fruit"))
```

$$
P(Burgundy | Fruit) = \frac{P(Burgundy \space\&\space Fruit)}{p(Fruit)}
$$

```{r}
fruit <- nrow(filter(wino, fruit))/nrow(wino)
burg_n_fruit <- nrow(filter(wino, province=="Burgundy" & fruit))/nrow(wino)
print(burg_n_fruit/fruit)
```

# 3. Naive Bayes Algorithm

We train a naive bayes algorithm to classify a wine's province using: 1. An 80-20 train-test split. 2. Three features engineered from the description 3. 5-fold cross validation.

We report Kappa after using the model to predict provinces in the holdout sample.

```{r, Feature Engineering 2}
winos <- wino %>% 
  mutate(yearf = as.factor(year)) %>% 
  mutate(spicy = str_detect(description,"spicy")) %>% 
  mutate(tart = str_detect(description,"tart")) %>%
  mutate(pink = str_detect(description,"pink")) %>%
  select(-description, year)

glimpse(winos)
```

```{r, Train-test Split}
set.seed(505)
winos_index <- createDataPartition(winos$province, p = 0.80, list = FALSE)
train <- wino[ winos_index, ]
test <- wino[-winos_index, ]
```

```{r, Training w/5-fold Cross Validation}
fit <- train(province ~ .,
             data = train, 
             method = "naive_bayes",
             metric = "Kappa",
             trControl = trainControl(number = 5, method = "cv"))
fit
```


# 4. Frequency Differences

We find the three words that most distinguish New York Pinots from all other Pinots.

```{r}
nydf <- wine %>%
  mutate(province = ifelse(province == "New_York", "NY", "Not_NY"))

df <- nydf %>%
  unnest_tokens(word, description) %>% 
  anti_join(stop_words) %>% 
  filter(str_detect(string = word, pattern = "[a-z+]")) %>%  
  filter(str_length(word)>3) %>%
  group_by(word) %>% 
  mutate(total=n()) %>% 
  ungroup()
head(df)

newdf <- df%>% 
    filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling"))) %>% 
    filter(total > 400) %>% 
    group_by(province, word) %>%
    count() %>% 
    group_by(province) %>%
    mutate(proportion = n / sum(n)) %>% 
    pivot_wider(id_cols = word, names_from = province, values_from = proportion)
head(newdf)

newdf%>%
    ggplot(aes(x = NY, y = Not_NY, color = abs(NY - Not_NY))) +
      geom_abline(color = "gray40", lty = 2) +
      geom_jitter(alpha = 0.1, size = 2.5, width = 0.3, height = 0.3) +
      geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
      scale_x_log10(labels = percent_format()) +
      scale_y_log10(labels = percent_format()) +
      scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
      theme(legend.position="none") +
      labs(x = "NY", y = "Not NY", title = "Words describing Pinot Noir from NY and Not NY")
```

```{r}
dtxt <- df %>% 
  filter(!(word %in% c("wine","pinot","drink","noir","vineyard","palate","notes","flavors","bottling","bottle","finish"))) %>% 
  filter(total > 400) %>% 
  group_by(province, word) %>%
  count() %>% 
  group_by(province) %>% 
  mutate(proportion = n / sum(n)) %>% 
  pivot_wider(id_cols = word, names_from = province, values_from = proportion) %>% 
  mutate(diff=NY-Not_NY) 

dtxt %>%
  top_n(25, diff) %>%
  mutate(word = reorder(word, diff)) %>%
  ggplot(aes(word, diff)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

**Answer:** The three most popular words in New York pinots are "cherry", "tannins","concentrated" located on the top of the bar graph. 